---
title: "Innovate UK Innovation Loan — Financial Analysis"
subtitle: "EOI: 9 October 2026 · £100k–£5M · 7.4% p.a. (3.7% + 3.7% deferred)"
author: "jessiorg / innovate-uk-innovation-loan"
date: today
format:
  html:
    self-contained: true
    code-fold: show
    code-tools: true
    toc: true
    toc-depth: 3
  pdf:
    toc: true
---

# 1. Loan Configuration

```{r}
#| label: config
#| include: false

CFG <- list(
  loan_amount     = 500000,
  project_years   = 2,
  extension_years  = 0,
  repayment_years  = 5,
  rate_drawdown   = 0.037,
  rate_repayment  = 0.074
)

n_proj  <- CFG$project_years   * 4
n_ext   <- CFG$extension_years * 4
n_repay <- CFG$repayment_years * 4
r_q_d   <- CFG$rate_drawdown  / 4
r_q_r   <- CFG$rate_repayment / 4
```

Configuration:

- **Loan amount:** £`r format(CFG$loan_amount, big.mark = ",")`
- **Project period:** `r CFG$project_years` years (`r n_proj` quarters)
- **Repayment period:** `r CFG$repayment_years` years (`r n_repay` quarters)
- **Drawdown rate:** `r scales::percent(CFG$rate_drawdown)` per year (`r scales::percent(r_q_d)` per quarter)
- **Repayment rate:** `r scales::percent(CFG$rate_repayment)` per year

---

# 2. Draw-Down Schedule

During the project period, the full loan (or tranche) is drawn in the first quarter. Interest at **3.7% is payable quarterly**; an equal amount is **deferred** to the repayment phase.

```{r}
#| label: drawdown

quarter     <- 1:n_proj
drawn       <- ifelse(quarter == 1, CFG$loan_amount, 0)
int_pay     <- drawn * r_q_d
int_def     <- drawn * r_q_d
cum_drawn   <- cumsum(drawn)
cum_def     <- cumsum(int_def)

drawdown <- data.frame(
  Quarter               = quarter,
  Drawn                 = drawn,
  `Interest.payable`    = int_pay,
  `Interest.deferred`   = int_def,
  `Cumulative.drawn`    = cum_drawn,
  `Cumulative.deferred` = cum_def
)

drawdown
```

::: {.callout-note}
**Note on staged drawdowns:** The model draws the full amount in Q1. If you stage your drawdown across quarters (e.g. £200k Q1, £300k Q3), deferred interest accrues on a smaller balance — reducing total cost. Adjust the `drawn` vector in the R code to model your actual drawdown plan.
:::

---

# 3. Loan Summary

```{r}
#| label: summary

principal     <- CFG$loan_amount
def_total     <- tail(cum_def, 1)
total_repay   <- principal + def_total
q_payment     <- total_repay * r_q_r / (1 - (1 + r_q_r)^(-n_repay))
total_interest<- q_payment * n_repay - principal
```

| Item | Value |
|---|---|
| Principal borrowed | £`r format(principal, big.mark = ",")` |
| Deferred interest (project phase) | £`r format(def_total, big.mark = ",")` |
| **Total to repay** | **£`r format(total_repay, big.mark = ",")`** |
| Quarterly payment (repayment phase) | £`r format(q_payment, big.mark = ",")` |
| Total interest paid | £`r format(total_interest, big.mark = ",")` |
| Effective cost of borrowing | `r scales::percent(total_interest / principal)` of principal |

---

# 4. Amortisation — Repayment Period

Quarterly annuity at **7.4%** on the full amount outstanding (principal + deferred interest).

```{r}
#| label: amort

opening  <- numeric(n_repay)
interest <- numeric(n_repay)
capital  <- numeric(n_repay)
closing  <- numeric(n_repay)

opening[1] <- total_repay
for (i in 1:n_repay) {
  interest[i] <- opening[i] * r_q_r
  capital[i]  <- q_payment - interest[i]
  closing[i]  <- opening[i] - capital[i]
  if (i < n_repay) opening[i + 1] <- closing[i]
}

amort <- data.frame(
  Quarter          = 1:n_repay,
  Opening.Balance  = round(opening, 2),
  Interest         = round(interest, 2),
  Capital          = round(capital, 2),
  Closing.Balance  = round(closing, 2)
)

rbind(head(amort, 4), tail(amort, 2))
```

*Showing first 4 and last 2 quarters only.*

---

# 5. Full Cashflow Timeline

```{r}
#| label: cashflow

n_total   <- n_proj + n_repay
phase     <- c(rep("Project", n_proj), rep("Repayment", n_repay))
drawdown_out <- c(drawdown$Drawn, rep(0, n_repay))
interest_out <- c(drawdown$`Interest.payable`, rep(0, n_repay))
capital_out  <- c(rep(0, n_proj), capital)
net_cf       <- drawdown_out - interest_out - capital_out
cum_cf       <- cumsum(net_cf)

full_cf <- data.frame(
  Quarter      = 1:n_total,
  Phase        = phase,
  Drawdown     = drawdown_out,
  Interest.Out  = interest_out,
  Capital.Out  = capital_out,
  Net.Cashflow = round(net_cf, 2),
  Cum.Cashflow = round(cum_cf, 2)
)

full_cf
```

---

# 6. DSCR Covenant Test

The Debt Service Coverage Ratio must be **≥ 1.2x** during the repayment period. Update the `revenue` vector in the R code block to match your business plan.

```{r}
#| label: dscr

revenue  <- c(100000, 250000, 500000, 750000, 1000000)  # ← edit to your plan
margin   <- 0.20                                           # ← edit EBITDA margin
ann_debt <- q_payment * 4

dscr_df <- data.frame(
  Year         = 1:5,
  Revenue      = revenue,
  EBITDA       = revenue * margin,
  Debt.Service = round(ann_debt, 0),
  DSCR         = round((revenue * margin) / ann_debt, 2),
  Covenant     = ifelse((revenue * margin) / ann_debt >= 1.2, "PASS", "FAIL")
)

dscr_df
```

**Minimum: 1.2x DSCR · 1.1x liquidity ratio** (tested quarterly via management accounts)

::: {.callout-warning}
At a £500k loan with a 5-year repayment, the annual debt service is ~£121,665. At 20% EBITDA margin you need **≥ £608k annual revenue** to clear the 1.2x covenant. Pre-revenue businesses must show a credible path to this threshold within the loan term.
:::

---

# 7. Security & Covenants for Pre-Revenue Businesses

## The Debenture — What It Actually Means

Innovate UK takes a **mortgage debenture over your company assets**. This is a fixed charge, not a floating charge, registered at Companies House. It covers:

- Equipment and plant
- IP (patents, trademarks, know-how)
- Trade receivables
- Inventory

**Critical for pre-revenue companies:** IP and intangible assets do **not** need to be independently valued. The loan is not secured against the assessed value of your IP — it is secured against the company's asset base. IUK assesses the loan based on the **quality of the innovation and commercial potential**, not the balance-sheet value of assets.

## No Personal Guarantees

Directors, founders, and shareholders are **not required to give personal guarantees**. Your home and personal assets are not at risk.

## What IUK Actually Assesses for Pre-Revenue Companies

| Factor | What they look at |
|---|---|
| Innovation quality | TRL 6-7+, novelty, competitive moat |
| Commercial potential | Route to market, customer evidence (LOIs, pilots), revenue model |
| Management team | Track record, R&D delivery experience |
| Financial robustness | Cash runway, burn rate, how you service interest during the project |
| Funding gap | Why the loan is needed vs. equity or commercial debt |
| Repayment path | How you plan to repay — IPO, trade sale, revenue, refinancing |

## What Pre-Revenue Companies Must Demonstrate

1. **You cannot get equivalent funding from private sources** — this is the market failure IUK exists to address
2. **You have a credible route to revenue** — LOIs, pilot agreements, regulatory approvals in place
3. **You can service interest during the project** — at minimum, you need cash to cover the 3.7% payable interest
4. **Stress-tested cash flow** — what happens if the project is delayed 6 months?

## Security Options & Alternatives

Since there are no personal guarantees and IP is not independently valued, the practical security levers are:

| Option | Detail |
|---|---|
| **Debenture over assets** | Standard. Register fixed charge at Companies House. |
| **Co-investment** | Having equity co-investors (angels, VCs) in alongside reduces IUK risk perception significantly |
| **R&D tax credit offsets** | If you have a HMRC R&D tax credit claim, this can serve as de facto security |
| **Pledge of IP** | IP can be included in the debenture; IUK doesn't require a valuation but wants to see it included |
| **Credit support from a third party** | A parent company or shareholder loan subordinated to IUK's position |

## The Covenant Risk for Pre-Revenue Businesses

The covenants (liquidity 1.1x, DSCR 1.2x) are only tested during the **repayment period** — not during the project period. This gives you up to 3 years (project) + 2 years (extension) = **5 years of runway** before the DSCR covenant is enforced.

That is your window to get to revenue. If you cannot demonstrate a credible path to the revenue threshold by project completion, the loan will be harder to structure.

---

# 8. EOI — What to Prepare

## The 1,000-word Project Summary Must Cover

1. **Vision** — The problem, who has it, why it matters
2. **Innovation** — What is novel, TRL level, why incumbents cannot do this
3. **Market** — Addressable market, first customers, route to market
4. **Team** — Key people, track record, why they are right for this
5. **Funding gap** — Why a loan, why not equity or grant
6. **UK economic impact** — Jobs, exports, IP, supply chain; which Industrial Strategy sector

## The 11 EOI Questions

**8 tick-box** — loan amount, sector, R&D stage, prior IUK funding, co-investment status, interest deferral, SME status, no double funding

**3 unscored** — commercial maturity, cash position, prior IUK funding history (self-declared)

**2 written responses:**
- *Fundraising plan* — broader funding strategy, why the innovation loan is the right instrument now
- *Investor stakes* — existing investor commitments, how the loan fits alongside them

---

*Repository: [github.com/jessiorg/innovate-uk-innovation-loan](https://github.com/jessiorg/innovate-uk-innovation-loan) · EOI Deadline: 9 October 2026, 11:00am*
